\bf A hierarchical failure-time model for heterogenous field data

### 2018-01-18 paper/m-lb-m-technometrics.tex
Make Meeker's suggested changes

\if0\blind
{
  \title{\bf A hierarchical failure-time model for observational data exhibiting infant-mortality and wearout failure modes}
  \author{Eric Mittman\thanks{Department of Statistics, Iowa State University}\\
    and \\
    Colin Lewis-Beck\footnotemark[1]\\
    and \\
    William Q. Meeker\footnotemark[1]}
  \maketitle
} \fi

@@

\begin{abstract}
When analyzing field data on consumer products, model-based approaches to inference require a model with sufficient flexibility to account for multiple kinds of failure. The causes of failure, while not interesting to the consumer per se, can lead to various observed lifetime distributions. Because of this, standard lifetime models, such as Weibull or lognormal may be inadequate. Usually cause-of-failure information will not be available to the consumer and thus traditional competing risk analyses cannot be performed.
Furthermore, when the information carried by lifetime data are limited by sample size, censoring and truncation, estimates can be unstable and suffer from imprecision. These limitations are typical; for example, lifetime data for high-reliability products will naturally tend to be right-censored.

In this paper we present a method for joint estimation of multiple lifetime distributions based on the Generalized Limited Failure Population (GLFP) model. This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  early failures due to ``infant mortality'' and failures due to wearout. We fit the GLFP model using a hierarchical modeling approach.  Borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.  Moreover, using our Bayesian method, comparison of different product brands is straightforward because estimation of arbitrary functionals is easy using draws from the joint posterior distribution of the model parameters. Potential applications include assessment and comparison of reliability to inform purchasing decisions.

@@

The reliability of many engineered products follows a similar pattern. Relatively high rates of failure occurring early (``infant mortality'') is due to manufacturing defects.  After this ``burn-in'' period, failure rates stabilize after the majority of defective units have failed.  Finally, after prolonged use, rates of failure increase due to wearout.  Ignoring this pattern in failure rates can lead to spurious inferences about a product's reliability and suboptimal decisions. This highlights the importance of choosing a sufficiently flexible model.

@@

\begin{itemize}
\item Making purchase decisions from among different suppliers of the components or subsystems
\item Making predictions for the number of needed replacement components or subsystems for spare part provisioning
\end{itemize}

\subsection{Model considerations}
Nonparametric methods require relatively few assumptions, but may not be suitable for prediction when extrapolation is required. Moreover, nonparametric estimates made using small samples, or data which are heavily censored may be too noisy to be useful. Appropriate parametric models can make better use of available data. For inference, likelihood-based methods are able to deal with truncated and/or censored data. Furthermore, by assuming a parsimonious representation of the data-generating process, parametric models admit simpler, more intuitive ways to compare populations of interest, borrow strength across sub-populations, and incorporate prior information.

@@

The GLFP model has a meaningful parameterization and accommodates lifetime data with both infant mortality and wearout.  Unfortunately, it can require a lot of data to fit due to the model's complexity. When multiple sources of information are available, partial pooling, accomplished via hierarchical modeling, can reduce the amount of data required to produce stable parameter estimates. When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data. As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations which imposes ``soft'' constraints on model parameters. This enables estimation of the lifetime distribution for all populations via shrinkage toward a ``pooled" model, with the degree of shrinkage inversely related to the amount of population specific information. We demonstrate computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained via Markov Chain Monte Carlo (MCMC).

\subsection{Overview}
The structure of the paper is as follows. Section~\ref{sec:Background} introduces the proposed method and potential applications and examines previous related work.  Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates its use with a subset of hard drive failure data from the company, Backblaze.  In Section~\ref{sec:Data}, we describe the complete Backblaze Hard Drive data set, which is used throughout as a motivating example. In Section~\ref{sec:Hierarchical GLFP model} we discuss using hierarchical modeling of the GLFP model parameters to extend the GLFP model to multiple populations. Section~\ref{sec:Data analysis} applies the hierarchical GLFP model to the hard drive data. First, we do model selection among four models of increasing complexity, using approximate leave-one-out cross-validation. Next, for the selected model, we assess the model fit using a graphical comparison of replicated data sets generated from the posterior distribution to the observed data. Finally, we present an evaluative comparison of populations taking into account practical considerations related to product lifetime. Section~\ref{sec:Discussion} describes potential applications of our model and limitations for inference implied by the model assumptions.

\section{Background}
\label{sec:Background}
In engineering applications, a product can often fail from one out of a set of possible malfunctioning components.  For example, a computer system can break if the mother board, disc drive or power supply stop working.  Circuit boards (CB) can fail due to a manufacturing defect or later as a result of wearout of certain components.  The general name for such products is a series system where the lifetime of the product is the minimum failure time across different components or risks \citep[Chapter 5]{nelson}.  A common assumption in series systems is the time to failure for each risk is statistically independent.  Thus, the overall reliability of a unit is the product of all the risks.  Parameter estimation is straightforward if the cause of failure is known for each observation.  However, in many situations the cause of failure is unknown to the analyst. This is referred to in the literature as masking.\\

Previous papers have employed various assumptions and methods to model masked lifetime failure data.  When modeling computer system failures \citet{reiser} assumed each observed failure came from a known subset of failure modes, and estimation was performed using a Bayesian approach.  \cite{chan} labeled the cause of circuit board failures as infant mortality, unknown, or wearout based on the time of observed failures.  This helped identify parameters when using maximum likelihood (ML) estimation.  Extending Chan and Meeker's analysis, \citet{basu} performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any assumptions about the cause of failure.  \cite{berger} introduced the Poly-Weibull distribution where the failure time is the minimum of a several Weibull distributions, corresponding to different failure modes.  \cite{ranjan} considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions.  Treating the unknown failure modes as incomplete data, an expectation maximization (EM) algorithm providing ML estimates was used, in addition to Bayesian estimation.

@@

We will use an alternative parameterization
where $\alpha$ is replaced by the $p$ quantile $t_{p}=\alpha \left
[-\log(1-p)\right ]^{\sigma}$. Replacing $\alpha$ in (1) with the
$\alpha=t_{p}/[-\log(1-p)]^{\sigma}$ gives
\begin{align*}
\nonumber
\Pr(T \leq t|t_{p},\sigma ) = F(t|t_{p},\sigma)&=1-
\exp \left [-\left (\frac{t}{t_{p}/[-\log(1-p)]^{\sigma}} \right )^{1/ \sigma}
\right ]\\
\label{equation:reparameterized.weibull.cdf}
&=1 - \exp\left[\log(1-p)\left(\frac{t}{t_{p}}\right)^{1/\sigma}\right], \quad t > 0.
\end{align*}
There are two important reasons for using this parameterization.
\begin{itemize}
\item
Especially with a high-reliability product, it will be easier to
elicit prior information about a quantile  ($t_p$) in the lower tail of the
distribution than it will be to elicit prior information about
$\alpha$. In addition, there is generally available
information about the shape parameter, $\sigma$, for a given failure
mechanism (e.g., if the failure is due to a wearout mechanism, then
it is known that $\sigma<1$). 
\item
Because of heavy censoring in reliability field data, the parameter
estimates of the $\mu$ and $\sigma$ parameters will generally be
highly correlated and thus specification of independent marginal
prior distributions would be be inappropriate. On the other hand,
estimates of $t_{p}$ and $\sigma$, for some appropriately chosen
value of $p$, will be approximately independent, allowing the easier
elicitation and specification of independent marginal prior
distributions. For example, if a typical field data set has 10\% of units
failing, then choosing $t_{0.05}$ would 
work well.
\end{itemize}

@@

The GLFP model can be understood as a mixture model with a binary latent variable, $\delta_i\ind \op{Bernoulli}(\pi)$. $\delta_i$ is an indicator for whether or not unit $i$ is defective. Expressed conditional on~$\delta_i$,
\begin{align*}
P(T\le t | \delta_i=1) =& 1 -(1-F_1(t))(1-F_2(t))\\
P(T\le t | \delta_i=0) =& F_2(t).
\end{align*}

The parameter $\pi$ represents the proportion of units susceptible to early failure, hence susceptible to both failure modes. Here the cause of failure is not assumed to be known, thus units from the same population are exchangeable.

Taking the derivative of the (marginal) cdf, the density for the GLFP is
\begin{align*}
h(t|\pi, t_{p1},\sigma_1, t_{p2}, \sigma_2) = \pi & f_1(t|t_{p1},\sigma_1)\left(1-F_2(t|t_{p2},\sigma_2)\right) + \\ & f_2(t|t_{p2},\sigma_2)\left(1-\pi F_1(t|t_{p1},\sigma_1\right).
\end{align*}

@@

\section{Motivating example}
\label{sec:Data}
\subsection{Backblaze hard drive data}
Backblaze is a company that offers cloud backup storage to protect against data loss.  Since 2013 it has been collecting data on hard drives operating at its facility.  The purpose is to provide consumers and businesses with reliability information on different hard drive-models.  The hard drives continuously spin in controlled storage pods.  Drives are run until failure.  When a hard drive fails it is removed and replaced.  In addition, the number of storage pods is increasing as Backblaze adds drives to its storage capacity.  Every quarter Backblaze makes its hard drive data publicly available through their website (\url{https://www.backblaze.com/b2/hard-drive-test-data.html}, Accessed April 1, 2016)\nocite{backblaze}. In addition, Backblaze publishes summary statistics of the different drive-models currently operating.  No other analysis or modeling of the failure data is provided.  Backblaze does, however, encourage others to further analyze their data.

As of the first quarter of 2016, Backblaze was collecting and reporting data on 63 different drive-models.  Some drive-models have been running since 2013 or before, while others were added at a later date.  Data have been reported on 75,297 different hard drives that are or were in operation.  The number of drives varies by drive-model; some drive-models only have a service record for a single drive whereas the maximum number of daily service records for a single drive-model is 35,860.  Figure 1 shows a scatterplot of the total observed running time in hundred of thousands hours versus the total number of failures for drive-models with at least 3 failures.  For model identification, a minimum of three failures was the criterion for hard drive brands to be included in our analysis.

\begin{figure}[H]
  %\includegraphics[width=.9\textwidth]{data-sum.pdf}
  \includegraphics[width=.9\textwidth]{dm-summ-scatter.pdf}
  \caption{Scatterplot of total observed running time (hundred thousands of hours) versus total number of observed failures.  Each number labels a unique drive-model.   Axes are on the log scale.}
  \label{drive-scatter}
\end{figure}

\subsection{Analysis of a single drive-model}
\label{subsec:ex1}
To illustrate the GLFP model, we will present an analysis of the drive-model with the most observed failures, Drive-model 14.  The Kaplan-Meier (K-M) estimate plotted below suggests at least two failure modes.  The curve has a slow ramp up in failures till about 10,000 hours (early failures), and increases more rapidly from about 10,000 to 20,000 hours (presumably wearout failures).  In addition, computer hard drives are known to have a mixture of early and late failures \citep{chan}.  Therefore, there is both an empirical and a theoretical justification to apply the GLFP model.

@@

To estimate the parameters of the GLFP model we use a Bayesian approach, selecting proper prior distributions to improve identification of the model parameters, which also ensures a proper posterior distribution. We reparameterize the component Weibull distributions using the $0.50$ quantile for the infant mortality failure mode ($t_{0.5,1}$) and the $0.20$ quantile for the wearout failure mode ($t_{0.2,2}$). \\

When eliciting prior distribution information it is much easier to ask about recognizable characteristics of a distribution instead of the parameters of the distribution.  Following the approach used in \cite[Section 15.2.2]{intervals} we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions.  Using this convention the prior distributions used in our analyses are
%$$t_{0.5,1} \sim \op{Lognormal}(8,4), \quad t_{0.2,2} \sim \op{Lognormal}(10,4)$$
$$t_{0.5,1} \sim \op{Log-normal} \langle 1.7,7.6\times 10^6 \rangle,\quad
t_{0.2,2} \sim \op{Log-normal}\langle 8.6, 5.6\times 10^7 \rangle$$
$$\pi \sim \op{Logit-normal}\langle 1.0\times 10^{-3}, 7.1\times 10^{-1} \rangle,\quad
\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Log-normal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2 \rangle. $$
%\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$

@@

Table \ref{table:1} gives the posterior median and 95\% credible intervals for the 5 GLFP parameters for Drive-model 14.  The parameter $\pi$ is an estimate of the proportion of drives susceptible to infant mortality.  As expected, this proportion is small with a median value of 0.05.  The shape parameter estimates for the two Weibull distributions also match intuition.  The infant mortality failure mode puts posterior probability on a values of $\beta_1$ less than 1, which corresponds to a decreasing hazard.  Conversely, the credible interval for $\beta_2$, the wearout mode, is strictly above 1, implying an increasing hazard function.  The two quantiles are also well identified with the infant mortality failure mode having an much earlier time to reach the 0.50 quantile compared to the time to reach the 0.20 quantile for the wearout mode.

@@

In Figure \ref{ex1-overlay}, we overlay the posterior median of the fitted GLFP model onto the adjusted K-M estimate with axes on the Weibull probability scale.  The plot also contains 90\% pointwise uncertainty bands associated with each estimate. While the Weibull model is inadequate for these data, the GLFP model fits quite well, as it is able to adequately describe the rapid increase in the empirical cdf between 8000 and 20,000 hours. %The discrepancy around 30,000 suggests that the GLFP may overestimate the proportion failing in the far right tail of the distribution.  While this discrepancy may appear troubling, we suggest that the right tail is of less importance for electronic goods due to obsolescence.

@@

In order to describe an entire product population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model sub-population-specific parameters hierarchically, borrowing strength across sub-populations.  Let
\begin{equation}
T_{ig} \ind \op{GLFP}\left( \pi_g, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g} \right),
\end{equation}
where $g=1,\ldots,G$ indexes the sub-populations.  The likelihood for the hierarchical GLFP is a function of the sets of parameters $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$, one set for each sub-population, $g$.  Assuming the lifetimes of all units are independent within and across sub-populations, and conditional on the fixed parameters, the likelihood for the data, $i=1,\dots,n_g$,  is given by
\begin{equation*}
L(\bm{\Theta})= \prod_{g=1}^{G} \prod_{i=1}^{n_{g}} \left[\frac{h(t_{ig};\bm{\theta_g})}{1-H(t_{ig}^L;\bm{\theta_g})}\right]^{1-c_{ig}} \left[ \frac{1-H(t_{ig};\bm{\theta_g})}{1-H(t_{ig}^L;\bm{\theta_g})} \right]^{c_{ig}}
\end{equation*}
where $T_{ig}$ is the observed failure or survival time of unit $i$ in sub-population $g$; $t_{ig}^L$ is the left truncation time for unit $i$ in sub-population $g$; and $c_{ig}$ is an indicator if unit $i$ in sub-population $g$ is right censored.  All times are again in hours.
\subsection{Parameter distributions}

In a hierarchical model, the parameters are modeled as random variables, varying across different sub-populations.  For example
\begin{equation*}
\sigma_{1g} \ind \op{Log-normal} \left( \eta_{\sigma_1}, \tau^2_{\sigma_1} \right)
\end{equation*}
\begin{equation*}
\sigma_{2g} \ind \op{Log-normal} \left( \eta_{\sigma_2}, \tau^2_{\sigma_2}\right)\op{Tr}\left(0, 1\right)
\end{equation*}
\begin{equation}
\label{eq:hier-model}
t_{p_{1}g} = \exp\left(\mu_{1g} + \sigma_{1g}\,\Phi^{-1}(p_1)\right)  \ind \op{Log-normal} \left(\eta_{t_{p_1}}, \tau^2_{t_{p_1}}\right)
\end{equation}
\begin{equation*}
t_{p_{2}g} = \exp\left(\mu_{2g} + \sigma_{2g}\,\Phi^{-1}(p_2)\right)  \ind \op{Log-normal} \left(\eta_{t_{p_2}}, \tau^2_{t_{p_2}}\right)
\end{equation*}
\begin{equation*}
\pi_g \ind \op{Logit-normal}(\eta_\pi, \tau^2_\pi).
\end{equation*}

@@

The set of model specifications is chosen based data, interpretation of the model, as well as estimation considerations.  sub-population-model-specific parameters for the the wearout failure mode $(t_{p_{2}g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across sub-populations in the right tails of the failure distribution.  Going from a common model for all sub-populations and gradually increasing the complexity of the model, Model 4 is the most flexible.  Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the wearout failure mode to vary by sub-population.

>>>>

\if0\blind
{
  \title{\bf A hierarchical failure-time model for heterogenous field data}
  \author{Eric Mittman\thanks{Department of Statistics, Iowa State University}\\
    and \\
    Colin Lewis-Beck\footnotemark[1]\\
    and \\
    William Q. Meeker\footnotemark[1]}
  \maketitle
} \fi

@@

\begin{abstract}
When analyzing field data on consumer products, model-based approaches to inference require a model with sufficient flexibility to account for multiple kinds of failures. The causes of failure, while not interesting to the consumer per se, can lead to various observed lifetime distributions. Because of this, standard lifetime models, such as Weibull or lognormal may be inadequate. Usually cause-of-failure information will not be available to the consumer and thus traditional competing risk analyses cannot be performed.
Furthermore, when the information carried by lifetime data are limited by sample size, censoring and truncation, estimates can be unstable and suffer from imprecision. These limitations are typical; for example, lifetime data for high-reliability products will naturally tend to be right-censored.

In this paper we present a method for joint estimation of multiple lifetime distributions based on the Generalized Limited Failure Population (GLFP) model. This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  early failures (sometimes referred to in the literature as ``infant mortality'') and failures due to wearout. We fit the GLFP model to a heterogenous population of devices using a hierarchical modeling approach.  Borrowing strength across sub-populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.  Moreover, using this Bayesian method, comparison of different product brands across the heterogenous population is straightforward because estimation of arbitrary functionals is easy using draws from the joint posterior distribution of the model parameters. Potential applications include assessment and comparison of reliability to inform purchasing decisions.

@@

The reliability of many engineered products follows a similar pattern. Relatively high rates of early failure are due to manufacturing defects.  After this ``burn-in'' period, failure rates stabilize after the majority of defective units have failed.  Finally, after prolonged use, rates of failure increase due to wearout.  Ignoring this pattern in failure rates can lead to spurious inferences about a product's reliability and suboptimal decisions. This highlights the importance of choosing a sufficiently flexible model.

@@

\begin{itemize}
\item Making purchase decisions from among different suppliers of the components or subsystems.
\item Making predictions for the number of needed replacement components or subsystems for spare part provisioning.
\end{itemize}

\subsection{Model considerations}
Nonparametric methods require relatively few assumptions, but may not be suitable for prediction when extrapolation is required. Moreover, nonparametric estimates made using small samples, or data which are heavily censored may be too noisy to be useful. Appropriate parametric models can make better use of available data. For inference, likelihood-based methods are able to deal with truncated and/or censored data. Furthermore, by assuming a parsimonious representation of the data-generating process, parametric models admit simpler, more intuitive ways to compare sub-populations of interest, borrow strength across sub-populations, and incorporate prior information.

@@

The GLFP model has a meaningful parameterization and accommodates lifetime data with both early and wearout failures.  Unfortunately, it can require a lot of data to fit due to the model's complexity. When multiple sources of information are available, partial pooling, accomplished via hierarchical modeling, can reduce the amount of data required to produce stable parameter estimates. When multiple sub-populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data. As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across sub-populations which imposes ``soft'' constraints on model parameters. This enables estimation of the lifetime distribution for all sub-populations via shrinkage toward a ``pooled" model, with the degree of shrinkage inversely related to the amount of sub-population specific information. We demonstrate the computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained via Markov Chain Monte Carlo (MCMC).

\subsection{Overview}
The structure of the paper is as follows. Section~\ref{sec:Background} introduces the proposed method and potential applications and examines previous related work.  Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates its use with a subset of hard drive failure data from the company, Backblaze.  In Section~\ref{sec:Data}, we describe the complete Backblaze Hard Drive data set, which is used throughout as a motivating example. In Section~\ref{sec:Hierarchical GLFP model} we discuss using hierarchical modeling of the GLFP model parameters to extend the GLFP model to multiple sub-populations. Section~\ref{sec:Data analysis} applies the hierarchical GLFP model to the hard drive data. First, we do model selection among four models of increasing complexity, using approximate leave-one-out cross-validation. Next, for the selected model, we assess the model fit using a graphical comparison of replicated data sets generated from the posterior distribution to the observed data. Finally, we present an evaluative comparison of sub-populations taking into account practical considerations related to product lifetime. Section~\ref{sec:Discussion} describes potential applications of our model and limitations for inference implied by the model assumptions.

\section{Background}
\label{sec:Background}
In engineering applications, a product can often fail from one out of a set of possible malfunctioning components.  For example, a computer system can break if the mother board, disc drive or power supply stop working.  Circuit boards (CB) can fail due to a manufacturing defect or later as a result of wearout of certain components.  The general name for such products is a series system where the lifetime of the product is the minimum failure time across different components or risks \citep[Chapter 5]{nelson}.  A common assumption in series system modeling is the time to failure for the different risks are statistically independent.  Thus, the overall reliability of a unit is the product of the reliability of each of the risks.  Then parameter estimation is straightforward if the cause of failure is known for each observation.  However, in many situations the cause of failure is unknown to the analyst. This is referred to in the literature as masking.

Previous papers have employed various assumptions and methods to model masked lifetime failure data.  When modeling computer system failures \citet{reiser} assumed each observed failure came from a known subset of failure modes, and estimation was performed using a Bayesian approach.  \cite{chan} labeled the cause of circuit board failures as early (due to defects), unknown, or wearout based on the time of observed failures.  This helped identify parameters when using maximum likelihood (ML) estimation.  Extending Chan and Meeker's analysis, \citet{basu} performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any assumptions about the causes of failure.  \cite{berger} introduced the Poly-Weibull distribution where the failure time is the minimum of a several Weibull distributions, corresponding to different failure modes.  \cite{ranjan} considered a competing risk model for early and wearout failures as a mixture of Weibull and exponential distributions.  Treating the unknown failure modes as incomplete data, an expectation maximization (EM) algorithm providing ML estimates was used, in addition to Bayesian estimation.

@@

We will use an alternative parameterization
where $\alpha$ is replaced by the $p$ quantile $t_{p}=\alpha \left
[-\log(1-p)\right ]^{\sigma}$. Replacing $\alpha$ in (1) with the
$\alpha=t_{p}/[-\log(1-p)]^{\sigma}$ gives
\begin{align*}
\nonumber
\Pr(T \leq t|t_{p},\sigma ) = F(t|t_{p},\sigma)&=1-
\exp \left [-\left (\frac{t}{t_{p}/[-\log(1-p)]^{\sigma}} \right )^{1/ \sigma}
\right ]\\
\label{equation:reparameterized.weibull.cdf}
&=1 - \exp\left[\log(1-p)\left(\frac{t}{t_{p}}\right)^{1/\sigma}\right], \quad t > 0.
\end{align*}
There are two important reasons for using this parameterization.
\begin{itemize}
\item
Especially with a high-reliability product, it will be easier to
elicit prior information about a quantile  ($t_p$) in the lower tail of the
distribution than it will be to elicit prior information about
$\alpha$ (approximately the $0.63$ quantile). In addition, there is generally available
information about the shape parameter, $\sigma$, for a given failure
mechanism (e.g., if the failure is due to a wearout mechanism, then
it is known that $\sigma<1$). 
\item
Because of heavy censoring in reliability field data, the parameter
estimates of the $\mu$ and $\sigma$ parameters will generally be
highly correlated and thus specification of independent marginal
prior distributions would be be inappropriate. On the other hand,
estimates of $t_{p}$ and $\sigma$, for some appropriately chosen
value of $p$, will be approximately independent, allowing the easier
elicitation and specification of independent marginal prior
distributions. For example, if a typical field data set has 10\% of units
failing, then choosing $t_{0.05}$ would 
work well.
\item Bayesian MCMC estimation will be better behaved due to reduced correlation between $t_p$ and $\sigma$ (relative to $\alpha$ and $\sigma$).
\end{itemize}

@@

The GLFP model can be understood as a mixture model with a binary latent variable, $\delta_i\ind \op{Bernoulli}(\pi)$. $\delta_i$ is an indicator for whether unit $i$ is defective or not (i.e., susceptible to an early failure). Expressed conditional on~$\delta_i$,
\begin{align*}
P(T\le t | \delta_i=1) =& 1 -(1-F_1(t))(1-F_2(t))\\
P(T\le t | \delta_i=0) =& F_2(t).
\end{align*}

The parameter $\pi$ represents the proportion of units susceptible to early failure, and hence susceptible to both failure modes. Here the cause of failure is not assumed to be known, thus units from the same population are exchangeable.

Taking the derivative of the (marginal) cdf, the density for the GLFP model is
\begin{align*}
h(t|\pi, t_{p1},\sigma_1, t_{p2}, \sigma_2) = \pi & f_1(t|t_{p1},\sigma_1)\left(1-F_2(t|t_{p2},\sigma_2)\right) + \\ & f_2(t|t_{p2},\sigma_2)\left(1-\pi F_1(t|t_{p1},\sigma_1\right).
\end{align*}

@@

\section{Motivating example}
\label{sec:Data}
\subsection{Backblaze hard drive data}
Backblaze is a company that offers cloud backup storage to protect against data loss.  Since 2013 it has been collecting data on hard drives operating at its facility.  The purpose is to provide consumers and businesses with reliability information on different hard drive-models.  The hard drives continuously spin in controlled storage pods.  Drives are run until failure.  When a hard drive fails it is removed and replaced.  In addition, the number of storage pods is increasing as Backblaze adds drives to its storage capacity.  Every quarter Backblaze makes its hard drive data publicly available through their website (\url{https://www.backblaze.com/b2/hard-drive-test-data.html}, accessed January 18, 2018)\nocite{backblaze}. In addition, Backblaze publishes summary statistics of the different drive-models currently operating.  No other analysis or modeling of the failure data is provided, other than simple rates and proportions.  Backblaze does, however, encourage others to further analyze their data.

As of the first quarter of 2016, Backblaze was collecting and reporting data on 63 different drive-models.  Some drive-models have been running since 2013 or before, while others were added at a later date.  Data have been reported on 75,297 different hard drives that are or were in operation.  The number of drives varies by drive-model; some drive-models only have a service record for a single drive whereas the maximum number of daily service records for a single drive-model is 35,860.  Figure 1 shows a scatterplot of the total observed running time in hundred of thousands hours versus the total number of failures for drive-models with at least 3 failures.  For model identification, a minimum of three failures was the criterion for inclusion of drive-models into our analysis.

\begin{figure}[H]
  %\includegraphics[width=.9\textwidth]{data-sum.pdf}
  \includegraphics[width=.9\textwidth]{dm-summ-scatter.pdf}
  \caption{Scatterplot of total observed running time (hundred thousands of hours) versus total number of observed failures.  Each number labels a unique drive-model, numbered between 1 and 44.   Axes are on the log scale.}
  \label{drive-scatter}
\end{figure}

\subsection{Analysis of a single drive-model}
\label{subsec:ex1}
To illustrate the GLFP model, we will present an analysis of the drive-model with the most observed failures, Drive-model 14.  The Kaplan-Meier (K-M) estimate plotted below suggests at least two failure modes.  The curve has a slow ramp up in failures till about 10,000 hours (early failures), and increases more rapidly from about 10,000 to 20,000 hours (presumably wearout failures).  Besides our empirical observation, it is well known that computer hard drives have a mixture of early and late failures \citep{chan}.  Therefore, there is both an empirical and a theoretical justification to apply the GLFP model.

@@

To estimate the parameters of the GLFP model we use a Bayesian approach, selecting proper, but generally diffuse, prior distributions to improve identification of the model parameters. Having a proper prior distribution also ensures a proper posterior distribution. We reparameterize the component Weibull distributions using the $0.50$ quantile for the early failure mode ($t_{0.5,1}$) and the $0.20$ quantile for the wearout failure mode ($t_{0.2,2}$). \\

When eliciting prior distribution information it is much easier to ask about recognizable characteristics of a distribution instead of the parameters of the distribution.  Following the approach used in \cite[Section 15.2.2]{intervals} we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions.  Using this convention the prior distributions used in our analyses are
%$$t_{0.5,1} \sim \op{Lognormal}(8,4), \quad t_{0.2,2} \sim \op{Lognormal}(10,4)$$
$$t_{0.5,1} \sim \op{Lognormal} \langle 1.7,\,7.6\times 10^6 \rangle,\quad
t_{0.2,2} \sim \op{Lognormal}\langle 8.6, \, 5.6\times 10^7 \rangle$$
$$\pi \sim \op{Logit-normal}\langle 1.0\times 10^{-3},\, 7.1\times 10^{-1} \rangle,\quad
\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,\,1.3 \times 10^2 \rangle. $$
%\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$

@@

Table \ref{table:1} gives the posterior median and 95\% credible intervals for the 5 GLFP parameters for Drive-model 14.  The parameter $\pi$ is an estimate of the proportion of drives susceptible to early failure.  As expected, this proportion is small with a median value of 0.05.  The shape parameter estimates for the two Weibull distributions also match intuition.  The early failure mode puts posterior probability on a values of $\beta_1$ less than 1, which corresponds to a decreasing hazard.  Conversely, the credible interval for $\beta_2$, the wearout mode, is strictly above 1, implying an increasing hazard function.  The two quantiles are also well identified with the early failure mode having an much earlier time to reach the 0.50 quantile compared to the time to reach the 0.20 quantile for the wearout mode.

@@

In Figure \ref{ex1-overlay}, we overlay the posterior median of the fitted GLFP model onto the adjusted K-M estimate with axes on the Weibull probability scale.  The plot also contains 90\% pointwise credible bands associated with each estimate. While the Weibull model is inadequate for these data, the GLFP model fits quite well, as it is able to adequately describe the rapid increase in the empirical cdf between 8000 and 20,000 hours. %The discrepancy around 30,000 suggests that the GLFP may overestimate the proportion failing in the far right tail of the distribution.  While this discrepancy may appear troubling, we suggest that the right tail is of less importance for electronic goods due to obsolescence.

@@

In order to describe an entire product population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model sub-population-specific parameters hierarchically, borrowing strength across sub-populations.  Let
\begin{equation}
T_{ig} \ind \op{GLFP}\left( \pi_g, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g} \right),
\end{equation}
where $g=1,\ldots,G$ indexes the sub-populations.  The likelihood for the hierarchical GLFP model is a function of the sets of parameters $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$, one set for each sub-population, $g$.  Assuming the lifetimes of all units are independent within and across sub-populations, and conditional on the fixed parameters, the likelihood for the data, $i=1,\dots,n_g$,  is given by
\begin{equation*}
L(\bm{\Theta})= \prod_{g=1}^{G} \prod_{i=1}^{n_{g}} \left[\frac{h(t_{ig};\bm{\theta_g})}{1-H(t_{ig}^L;\bm{\theta_g})}\right]^{1-c_{ig}} \left[ \frac{1-H(t_{ig};\bm{\theta_g})}{1-H(t_{ig}^L;\bm{\theta_g})} \right]^{c_{ig}}
\end{equation*}
where $T_{ig}$ is the observed failure or survival time of unit $i$ in sub-population $g$; $t_{ig}^L$ is the left truncation time for unit $i$ in sub-population $g$; and $c_{ig}$ is an indicator if unit $i$ in sub-population $g$ is right censored.  All times are again in hours.
\subsection{Parameter distributions}

In a hierarchical model, the parameters are modeled as random variables, varying across different sub-populations.  For example
\begin{equation*}
\sigma_{1g} \ind \op{Lognormal} \left( \eta_{\sigma_1}, \tau^2_{\sigma_1} \right)
\end{equation*}
\begin{equation*}
\sigma_{2g} \ind \op{Lognormal} \left( \eta_{\sigma_2}, \tau^2_{\sigma_2}\right)\op{Tr}\left(0, 1\right)
\end{equation*}
\begin{equation}
\label{eq:hier-model}
t_{p_{1}g} = \exp\left(\mu_{1g} + \sigma_{1g}\,\Phi^{-1}(p_1)\right)  \ind \op{Lognormal} \left(\eta_{t_{p_1}}, \tau^2_{t_{p_1}}\right)
\end{equation}
\begin{equation*}
t_{p_{2}g} = \exp\left(\mu_{2g} + \sigma_{2g}\,\Phi^{-1}(p_2)\right)  \ind \op{Lognormal} \left(\eta_{t_{p_2}}, \tau^2_{t_{p_2}}\right)
\end{equation*}
\begin{equation*}
\pi_g \ind \op{Logit-normal}(\eta_\pi, \tau^2_\pi).
\end{equation*}

@@

This set of model specifications is chosen based on the data, interpretation of the model, as well as estimation considerations.  Sub-population-model-specific parameters for the the wearout failure mode $(t_{p_{2}g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across sub-populations in the right tails of the failure distribution.  Going from a common model for all sub-populations and gradually increasing the complexity of the model, Model 4 is the most flexible.  Model 4 allows for the probability of early failure as well as the shape and scale parameters for the wearout failure mode to vary across sub-populations.

For all of the nested models, the parameters for the early failure mode are common across sub-populations.  We found that there was often insufficient information to model these parameters hierarchically.  Moreover, assuming a common distribution for early failures provides a meaningful interpretation and comparison of $\pi_g$ and $\pi_{g'}$ ($g \neq g'$). Also the failure-time distribution of the defective subpopulation (as opposed to the proportion defective) is not of high interest.
